results

### 2022-04-12 code/analysis_code/exploration.Rmd
sunday updates
quad figure 1

>>>>

# stats for land use, p-value for land use
```{r}
ggplot(data = land_long, aes(x = value, y = particles_l)) + geom_point(size = .5) + facet_wrap(~name) + stat_cor(method = "spearman") + labs(x = "Total Area (sq. meters)", y = "Particles per Liter") + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

### 2022-04-12 code/analysis_code/particle_type.Rmd
quad figure 1

```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)
library(readxl) #for loading Excel files

@@

sum <- sum %>% mutate(fragment = fragment_avg_a + fragment_avg_b, 
                      fiber = fiber_avg_a + fiber_avg_b, 
                      bead = bead_avg_a + bead_avg_b, 
                      foam = foam_avg_a + foam_avg_b, 
                      pellet = pellet_avg_a + pellet_avg_b, 
                      film = film_avg_a + film_avg_b) %>% select(13:18)

@@

```{r}
part_sum %>% ggplot(aes(x = name, y = value, fill = name)) + 
  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5, position = position_dodge(width=0.9))) + ylim(0, 350) 
```

@@

>>>>

```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)
library(readxl) #for loading Excel files
library(ggpubr)

@@

sum <- sum %>% mutate(Fragment = fragment_avg_a + fragment_avg_b, 
                      Fiber = fiber_avg_a + fiber_avg_b, 
                      Bead = bead_avg_a + bead_avg_b, 
                      Foam = foam_avg_a + foam_avg_b, 
                      Pellet = pellet_avg_a + pellet_avg_b, 
                      Film = film_avg_a + film_avg_b) %>% select(13:18)

@@

```{r}
part_sum %>% ggplot(aes(x = name, y = value)) + 
  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5)) + ylim(0,350) + guides(fill=guide_legend("Particle Type")) + labs(y = "Microplastic Composition")
#position = position_dodge(width=0.9))) + ylim(0, 350) 
```

@@

#PARTICLE TYPE FIGURE
```{r}
fig_part_type <- part_sum %>% ggplot(aes(x = name, y = value)) + 
  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5), size = 3) + ylim(0,380) + guides(fill=guide_legend("Particle Type")) + labs(y = "No. of Particles", x = "") + theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1), axis.text.y=element_text(size=12))
plot(fig_part_type)

```

#PARTICLE COLOR FIGURE
```{r}
library(ggrepel)
# load particle color data
color <- read_excel(here("data", "raw_data", "mp_color.xlsx"))

# plot reordered stacked plot
color <- color %>%  mutate(color= str_to_title(color)) 
color <- color %>% rename(Color = color)
color$Color <- reorder(color$Color, color$number) #largest to smallest

# plot color data
fig_part_color <- color %>% ggplot(aes(x = Color, y = number, fill = Color)) + 
  geom_col() + geom_text(aes(label = percent(number/968), vjust = -0.5), size = 3) + ylim(0,460) + scale_fill_manual(values = c("Black" = "gray0", "Blue" = "royalblue2", "Green" = "seagreen2", "Other" = "orange1", "Red" = "red1", "TRN/White"= "gray61")) + labs(x = "", y = "Microplastic Composition") + labs(y = "No. of Particles")  + theme_bw() + theme(legend.position = "none")+ theme(axis.text.x = element_text(angle = 45, hjust=1))
plot(fig_part_color)
#stacked
#reorder

ggplot(data = color, aes(x = obs, y = number, fill = Color)) + geom_bar(stat = "identity", position = "fill") + scale_fill_manual(values = c("Black" = "gray0", "Blue" = "royalblue2", "Green" = "seagreen2", "Other" = "orange1", "Red" = "red1", "Transparent/White"= "gray61")) + labs(x = "", y = "Microplastic Composition") + guides(fill=guide_legend("Particle Color")) +
   geom_text(aes(label = percent(number/968)), size = 3, color = "white", position = position_fill(vjust = 0.5)) + theme_bw()+ theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
plot(fig_part_color)

```
# particle type by water body
```{r}
ggplot(data = part_sum, aes(x = body, y = value, fill = name)) + geom_bar(stat = "identity", position = "fill") + labs(x = "", y = "Microplastic Composition") + guides(fill=guide_legend("Particle Color")) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

### 2022-04-12 code/analysis_code/results.Rmd
quad figure 1

#load packages and data
```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)
library(gridExtra) #for savings images in grids
library(gridExtra) #for savings images in grids

@@

@@

P=value less than 0.05 - we conclude that there IS a significant difference between groups (though we don't yet know between what groups). From here, we may use the pairwise.wilcox.test() to calculate pairwise comparison
# pairwise wilcox test to see differences between groups
```{r}
pairwise.wilcox.test(all_counts_cv$particles_l, all_counts_cv$date,
                 p.adjust.method = "BH")
```

@@

#plot
body %>% ggplot(aes(x = body, y = particles_l, color = )) + 
  geom_boxplot()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Water Body", x = "Water Body", y = "Particles/Liter") + ylim(0,500) + scale_y_log10()

#kruskal wallis/mann whitney test - nonparametric test
kruskal.test(particles_l ~ body, data = body)
```

@@

This tells us that there is a significant difference between the MP concentrations in rivers and tributaries! Cool! We can try to tweak this boxplot figure to display those statistics, perhaps.

>>>>

#load packages and data
```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidyverse)
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)
library(gridExtra) #for savings images in grids
library(gridExtra) #for savings images in grids
library(ggpubr) #adding stats to figures

@@

#adding stats to graph
```{r}
# with line of best fit
ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() + geom_smooth(method = lm) + labs(x = "Particles per Liter", y = "Coefficient of Variance %") +
  stat_cor(method = "spearman") + scale_x_continuous(trans='log10') + theme_bw() # use log scale to better visualize all points

#no line of best fit
ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() + labs(x = "Particles per Liter", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') + theme_bw() # use log scale to better visualize all points
```

@@

P=value less than 0.05 - we conclude that there IS a significant difference between groups (though we don't yet know between what groups). From here, we may use the pairwise.wilcox.test() to calculate pairwise comparison
# pairwise wilcox test to see differences between groups
```{r}
pairwise.wilcox.test(all_counts_cv$particles_l, all_counts_cv$date,
                 p.adjust.method = "BH")

compare_means(particles_l ~ date, data = all_counts_cv, p.adjust.method = "BH")
```

@@

#rename
body$body <- fct_recode(body$body,
                                 "Main Rivers" = "river",
                                 "Tributaries" = "trib")
pv <- body %>% summarize(p.value = round(kruskal.test(particles_l ~ body)$p.value, digits = 3))

#plot
body %>% ggplot(aes(x = body, y = particles_l, color = )) + 
  geom_boxplot()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(x = "Water Body", y = "Particles per Liter") + ylim(0,500) + scale_y_log10() +
    geom_text(data=pv, aes(x=1.5, y=900, label=paste0("Kruskal-Wallis\n p=",p.value))) + theme_bw()

#kruskal wallis/mann whitney test - nonparametric test
kruskal.test(particles_l ~ body, data = body)
wilcox.test(particles_l ~ body, data = body)

summary(river)
summary(trib)
```

@@

This tells us that there is a significant difference between the MP concentrations in rivers and tributaries! Cool! We can try to tweak this boxplot figure to display those statistics, perhaps. 
# figure with stats for seasons
```{r}
compare_means(particles_l ~ date,  data = body)
my_comparisons <- list(c("nov_20", "feb_21"), c("nov_20", "apr_21"), c("nov_20", "july_21"), c("feb_21", "apr_21"), c("feb_21", "july_21"), c("apr_21", "july_21"))

fig_season <- body %>% ggplot(aes(x = date, y = particles_l, group = date)) + 
  geom_boxplot() + scale_color_discrete(name = "Water Body")+ 
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(x = "Sample Date", y = "Particles per Liter") + scale_x_discrete(labels=c("Fall 2020",
                            "Winter 2021",
                            "Spring 2021",
                            "Summer 2021"))+ theme_bw() + stat_compare_means(comparison=my_comparisons, p.adjust.method = "BH") + stat_compare_means(label.y = 5) + scale_y_log10()
plot(fig_season)
```
#arrange figures
```{r}
#map
#Load Libraries 
library(tidyverse)
library(viridis)
library(ggpubr)
#Load data
sites = readRDS(here::here("data","processed_data","mpwrfcombo.rds"))
sites$date = as.factor(sites$date)
sites$date = recode(sites$date, nov_20 = "Nov 2020", feb_21 = "Feb 2021", apr_21 = "April 2021", july_21 = "July 2021")
sites$date = ordered(sites$date, levels = c("Nov 2020", "Feb 2021", "April 2021", "July 2021"))
sites %>% ggplot(aes(x = date, y = site, fill= particles_l)) + 
  geom_tile() + 
  scale_fill_viridis() + 
  theme_bw() +
  ylab("Site") + 
  xlab("Collection Period") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = date, y = site, label = round(particles_l))) +
  labs(fill = "Particles per Liter")  
  #ggtitle("Seasonality of Microplastics Concentration")
sites %>% ggplot(aes(x = date, y = site, fill= log10(particles_l))) + 
  geom_tile() + 
  scale_fill_viridis() + 
  theme_bw() +
  ylab("Site") + 
  xlab("") + 
  scale_x_discrete(position = "top") +
  #theme(axis.text.x = element_text(angle = 45)) +
  geom_text(aes(x = date, y = site, label = round(log10(particles_l), 2))) +
  labs(fill = "Log10(Particles/Liter)") + 
  theme(legend.position = "bottom")
  #ggtitle("Seasonality of Microplastics Concentration")
hist_all = sites %>% ggplot(aes(x = particles_l)) + 
  geom_histogram() + 
  theme_bw() + 
  #facet_wrap(~date) +
  xlab("Particles per Liter") + 
  ylab("No. of Observations") +
  theme(axis.text.y=element_text(size=12))
hist_season = sites %>% ggplot(aes(x = particles_l)) + 
  geom_histogram() + 
  theme_bw() + 
  facet_wrap(~date) +
  xlab("Particles per Liter") + 
  ylab("No. of Observations") +
  theme(axis.text.y=element_text(size=12))
fig2 = ggarrange(hist_all, hist_season, ncol = 2)
plot(fig2)
sites %>% group_by(date) %>% tally()

#particle type = fig_part_type
#particle color = fig_part_color
quad1 <- ggarrange(hist_all, hist_season, fig_part_type, fig_part_color, nrow = 2, ncol = 2, labels = "AUTO", vjust = 1) 
#%>% ggexport(filename = here("results","manuscript_figures", "quad_1.png"))
plot(quad_1)

ggsave(filename = here("results","manuscript_figures", "quad_1.png")) #export resolution than ggexport
```
